### 2009-12-13 docs/marcello-dissertation/thesis/chapter10-appendix.tex
Reviews on the Appendix section
- Adding references to the execution logs and developed scripts
- Adding more description about the infrastructure setup
- Updating reference on chapter 2  (wrong)
- Updating the reference to the setup scripts on chapter 8.

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@701 559edde2-a4d4-44e0-96fc-4035e40d7396

@@

@@

@@

%TODO: add bootstrap message in case it is implemented

@@

\lstinputlisting[language=Bash,label=file:experiment-setup-executor,caption=Experiment
Main Execution Script]{../../../../netbeams/versions/v2/persistence/run-persistence-experiment}

\lstinputlisting[language=Python,label=file:experiment-setup-mongoshards,caption=Experiment
Main Execution Script]{../../../../netbeams/versions/v2/persistence/setup-mongo-shards.js}

\lstinputlisting[language=Bash,label=file:setup-os-shards-step1,caption=Experiment
Main Execution Script]{../../../../netbeams/versions/v2/persistence/setup-os-shards-step1}

\lstinputlisting[language=Bash,label=file:setup-os-shards-step2,caption=Experiment
Main Execution Script]{../../../../netbeams/versions/v2/persistence/setup-os-shards-step2}

\lstinputlisting[language=Python,label=file:experiment-export-python,caption=Data
Export to from mongoDB to OPeNDAP format.]{../../../../netbeams/versions/v2/persistence/data-export-python.py}

\lstinputlisting[language=Python,label=file:experiment-scenarios,caption=Experiment
Scenarios Implementation]{../../../../netbeams/versions/v2/persistence/experiment-client-scenarios.js}

\lstinputlisting[language=Java,label=file:main-ysi-data-handler,caption=The
DSP YSI Sonde Data Handler]{../../../../netbeams/versions/v2/apps/xml/jaxb/sonde/java-src/org/netbeams/dsp/ysi/SondeDataType.java}

\lstinputlisting[language=Java,label=file:random-ysi-data-generator,caption=Random
Test Data Generator]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPSondeProducer/src/org/netbeams/dsp/ysi/SondeTestData.java}

\lstinputlisting[language=Java,label=file:experiment-dsp-java-executor,caption=Inserting
Random Data to MongoDB]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/src/org/netbeams/dsp/persistence/controller/DSPMessageToMongoDBExperiment.java}

\subsection{Experiment Execution Output}

>>>>

The description of a distributed system can be seen next section.

@@

\subsection{Setting Up the Single and the Distributed Cluster Servers}
\label{sec:experiments-setup-single-distributed}
mongoDB can be used in two different ways: as a single server, using the single
process ``mongod'', or as a distributed cluster where each cluster node runs on
``mongod'' managed by a cluster head ``mongos''.

Before starting the main cluster head, all cluster nodes must be started.
Listing \ref{file:shard-setup-192.168.1.2.log} shows the manual setup process
performed in the Mac OS X 64bits Host. It starts two different shards, with
their own respective directory, and starts the process ``mongod'' for each of
them, forwarding the execution output to their respective log artifacts. Note
that the IP address is capture in the beginning of the process because it will
be used while setting up the cluster head.

In order to automate the manual process, different automation scripts were
written using the Linux shell bash execution, as well as the ones used to setup
mongoDB using Javascript, the language used by mongoDB shell client. First, the
scripts to setup the mongoDB distributed cluster is shown in Listing
\ref{file:setup-mongo-shards.js}. It enables shard support for the different
shards that must be already running, and then enable it on the collection
``SondeDataContainer'' for the database ``netbeams''. As it is shown in the
script, the declaration of the shards are executed in the ``admin'' database,
while shards are in the database ``config'', since it is the main cluster
metadata server.

The first script created was the automated shell bash script to setup local
shards directories and processes, as well as the cluster head metadata server
(config) and the start of the cluster head ``mongos''. Each local shard can be
candidates to store the collected data, depending on the decisions made by the
mongos cluster head. Listing \ref{file:setup-experiment-shards.sh} expects to be
executed by passing the number of local shards as a parameter to create the
local directories and start the processes described. Listing
\ref{file:shards-and-mongos-cluster-head-setup-execution.log} shows the output
of the exeuction of the setup script, incorporating the output of the Listing
\ref{file:setup-mongo-shards.js}. Note that the same IP address, shown in the
setup of the cluster nodes, is also shown in the list of the shards.

\subsection{Running the Experiments}

After setting up the infrastructure, the main experiment script can be executed
in the client-side. Considering the distributed server approach, the process
``mongos'' must be running in the server host. The infrastructure for the
execution of the experiments used Sun Microsystem's VirtualBox to run different
operating systems at the same time. The cluster head ``mongos'' was running on
a Ubuntu Linux 9.04 64 bits together with the cluster metadata server
``mongod'' referred to as ``config'' server. The cluster nodes were running on
the Mac OS X 64 bits on the host server, running two different shards. Finally,
the execution of the experiments used Ubuntu Linux 8.04 32 Bits.

As developed for the infrastructure setup, the execution of the experiments
were performed using automation scripts written in different languages.
First, the experiments for the ``Create" operations was written in Java
as shown in Listing \ref{file:experiment-dsp-java-executor}. Similarly,
remaining operations ``Retrieve, Update and Delete" were written in
Javascript, as shown in Listing \ref{file:experiment-scenarios}. Then, the
execution of this Java class is performed by the main experiment shell script
displayed in Listing \ref{file:experiment-setup-executor}. It requires the
number of randomly created YSI Sonde data to be executed, as shown in Listing
\ref{cmd:run-persistence-experiment}.

@@

The first experiment conducted was running the main experiment script of Listing
\ref{file:experiment-setup-executor}, as one of the execution runs shown in
Listing \ref{file:experiment-run.log}. The total number of inserted documents
measured every 20 seconds is displayed. Then, the execution of the use cases
was performed, as shown in Listing \ref{file:URD-scenarios-execution.log}.
Additional execution logs are shown in Section
\ref{sec:experiment-execution-logs}.

@@

@@

\lstinputlisting[language=Bash,label=file:shard-setup-192.168.1.2.log,caption=Remote
Shards Setup Execution Log]{../../../../netbeams/versions/v2/persistence/logs/shard-setup-192.168.1.2.log}

\lstinputlisting[language=Bash,label=file:setup-experiment-shards.sh,caption=Remote
Shards Cluster Head and Metadata Setup Bash Script]{../../../../netbeams/versions/v2/persistence/setup-experiment-shards.sh}

\lstinputlisting[label=file:setup-mongo-shards.js,caption=Shards
Setup in Javascript]{../../../../netbeams/versions/v2/persistence/setup-mongo-shards.js}

\lstinputlisting[language=Bash,label=file:experiment-setup-executor,caption=Main Client
Experiment Execution Bash Script]{../../../../netbeams/versions/v2/persistence/run-persistence-experiment}

\lstinputlisting[label=file:experiment-scenarios,caption=Experiment
Scenarios Implementation in Javascript]{../../../../netbeams/versions/v2/persistence/experiment-client-scenarios.js}

\lstinputlisting[language=Python,label=file:experiment-export-python,caption=Data
Export to from mongoDB to OPeNDAP Format Python Script]{../../../../netbeams/versions/v2/persistence/data-export-python.py}

\lstinputlisting[language=Java,label=file:main-ysi-data-handler,caption=The
DSP YSI Sonde Data Java Handler]{../../../../netbeams/versions/v2/apps/xml/jaxb/sonde/java-src/org/netbeams/dsp/ysi/SondeDataType.java}

\lstinputlisting[language=Java,label=file:random-ysi-data-generator,caption=Random
Test Data Java Generator]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPSondeProducer/src/org/netbeams/dsp/ysi/SondeTestData.java}

\lstinputlisting[language=Java,label=file:experiment-dsp-java-executor,caption=Inserting
Random Data to MongoDB Java Experiment]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/src/org/netbeams/dsp/persistence/controller/DSPMessageToMongoDBExperiment.java}

\subsection{Experiment Execution Output Logs}
\label{sec:experiment-execution-logs}

\lstinputlisting[language=Bash,label=file:shards-and-mongos-cluster-head-setup-execution.log,caption=Execution
of Shards Setup Log]{../../../../netbeams/versions/v2/persistence/logs/shards-and-mongos-cluster-head-setup-execution.log}

\lstinputlisting[language=Bash,label=file:experiment-run.log,caption=Main Client
Experiment Execution Log]{../../../../netbeams/versions/v2/persistence/experiment-run.log}

\lstinputlisting[language=Bash,label=file:URD-scenarios-execution.log,caption=RUD
Scenarios Execution Log]{../../../../netbeams/versions/v2/persistence/logs/URD-scenarios-execution.log}

\lstinputlisting[language=Bash,label=file:mongos-shards-cluster-head-metadata-config-server-192.168.1.10:10000.log,caption=Remote
Shards Cluster Head Execution Log]{../../../../netbeams/versions/v2/persistence/logs/mongos-shards-cluster-head-metadata-config-server-192.168.1.10:10000.log}

\lstinputlisting[language=Bash,label=file:mongos-shards-cluster-head-192.168.1.10.log,caption=Remote
Shards Cluster Head Execution]{../../../../netbeams/versions/v2/persistence/logs/mongos-shards-cluster-head-192.168.1.10.log}

\lstinputlisting[language=Bash,label=file:shard-192.168.1.2:20002.log,caption=Remote
Shard Execution Log]{../../../../netbeams/versions/v2/persistence/logs/shard-192.168.1.2:20002.log}

### 2009-12-13 docs/marcello-dissertation/thesis/chapter2-state-of-the-art.tex
Reviews on the Appendix section
- Adding references to the execution logs and developed scripts
- Adding more description about the infrastructure setup
- Updating reference on chapter 2  (wrong)
- Updating the reference to the setup scripts on chapter 8.

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@701 559edde2-a4d4-44e0-96fc-4035e40d7396

Depending on the application of the collected sensor data, the location in
which the data was collected may be needed. For example, \cite{sn-geo-metadata}
shows the geographic location of the collected data is critical to identify
the locations of the events of the sensor device \cite{sn-ex02}. Similarly,
the position of any event captured by a black-box of an aircraft carries 
metadata for the event. Examples of this are the metadata event =
``high-pressure" of 1304, at the altitude of 1,500 feet above the ocean on GPS
coordinates of latitude 13.445 and longitude -23,003. This example clearly
gives necessary identification of where the event of high-pressure occurred.
The precision of the location can be important for the collected data because
it helps researchers correlate different measurements based on exact locations
of the data. For instance, \ref{sn-provenance} depicts an example where two
different sensor networks covering the same physical area: one covering the
reginal traffic and another snowfall. Although each of them has individual
functions, collectively they could allow real-time navigation service to answer
queries regarding expected traffic patterns based on the current snowfall and
advise drivers accordingly by reusing collected data from both sensor networks.

>>>>

Depending on the application of the collected sensor data, the location in
which the data was collected may be needed. For example, \cite{sn-geo-metadata}
shows the geographic location of the collected data is critical to identify
the locations of the events of the sensor device \cite{sn-ex02}. Similarly,
the position of any event captured by a black-box of an aircraft carries 
metadata for the event. Examples of this are the metadata event =
``high-pressure" of 1304, at the altitude of 1,500 feet above the ocean on GPS
coordinates of latitude 13.445 and longitude -23,003. This example clearly
gives necessary identification of where the event of high-pressure occurred.
The precision of the location can be important for the collected data because
it helps researchers correlate different measurements based on exact locations
of the data. For instance, \cite{sn-provenance} depicts an example where two
different sensor networks covering the same physical area: one covering the
reginal traffic and another snowfall. Although each of them has individual
functions, collectively they could allow real-time navigation service to answer
queries regarding expected traffic patterns based on the current snowfall and
advise drivers accordingly by reusing collected data from both sensor networks.

### 2009-12-13 docs/marcello-dissertation/thesis/chapter8-experimental-results-correct-behavior-and-performance.tex
Reviews on the Appendix section
- Adding references to the execution logs and developed scripts
- Adding more description about the infrastructure setup
- Updating reference on chapter 2  (wrong)
- Updating the reference to the setup scripts on chapter 8.

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@701 559edde2-a4d4-44e0-96fc-4035e40d7396

The simulation of both the single and distributed system used both a Mac OS 10
Snow Leopard 64bit, with a total of 8GB of physical memory. However, in order
to simulate different shards in different machines, the use of a VirtualBox
\cite{virtualization} provided the use of guest operating systems to run and
act as the mongoDB shards. In this way, two different versions of the Ubuntu
Linux were used: one 8.04 32bits and another 9.04 64bits. The configuration
used for the single server only uses the process ``mongod'' to be started,
while the clustered setup must be enabled in the database, as shown in
Listing \ref{file:experiment-setup-mongoshards}, Listings
\ref{file:setup-os-shards-step1} and \ref{file:setup-os-shards-step2}.

@@

\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.5]{../diagrams/taxonomy-data-model-modified}
  \caption{Data Model Taxonomy}
  \label{fig:taxonomy-data-model-modified}
\end{figure}

@@

Physical memory is limited and, for this reason, it may not be enough to run
the experiments for total workloads. In this case, the Java Virtual Memory was
totally consumed by one of the revisions of the experiments setup. Therefore, 
it was learned that large amount of data loads require different techniques to
manage data in memory accordingly, such as creating different object pools as
temporary caches for objects whose internal state are similar to others in the
Java Virtual Machine.

>>>>

The simulation of both the single and distributed system used both a Mac OS 10
Snow Leopard 64bit, with a total of 8GB of physical memory. However, in order
to simulate different shards in different machines, the use of a VirtualBox
\cite{virtualization} provided the use of guest operating systems to run and
act as the mongoDB shards. In this way, two different versions of the Ubuntu
Linux were used: one 8.04 32bits and another 9.04 64bits. The configuration
used for the single server only uses the process ``mongod'' to be started,
while the clustered setup must be enabled in the database, as in Section
\ref{sec:experiments-setup-single-distributed}.

@@

\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.65]{../diagrams/taxonomy-data-model-modified}
  \caption{Data Model Taxonomy}
  \label{fig:taxonomy-data-model-modified}
\end{figure}

@@

Physical memory is limited and, for this reason, it may not be enough to run
the experiments for total workloads. In this case, the Java Virtual Memory was
totally consumed by one of the revisions of the experiments setup. Therefore, 
it was learned that large amount of data loads require different techniques to
manage data in memory accordingly, such as creating different object pools as
temporary caches for objects whose internal state are similar to others in the
JVM. Also, operations that involve writing in disk, such as during the bulk
``Create'' operations, may lead to different database performance. First, the
experiment seemed to stop inserting objects in one shard while it ran for over
2 minutes without inserting any documents. The reason was that the partition
was ``/home'', completely full, leading to the problem. The mongos cluster
header did not realize that it was the time to switch to another shard.
Similarly, any global ``Retreave'' operation on a cluster blocks all the
shards. For this reason, dividing the write from the read load was necessary in
order to decouple these two operations.
